Fluent statistical computing interfaces for biological data analysis

### 2020-07-31 Rmd/05-embeddings.Rmd
spelling

We have shown in the case studies, that one to one linked brushing can
be used to identify distortions in the embedding, however we would
like extend this to one to many linked brushing, which would allow us to directly interrogate neighbourhood preservation. This form of brushing acts directly on a $k$-nearest neighbours ($k$-nn) graph computed from a
reference dataset: when a user brushes over a region in the embedding,
all the points that match the graphs edges are selected on the 
corresponding tour view. The reference data set for computing
nearest neighbours (for example a distance matrix, or the complete data matrix) 
can be independent of the tour or embedding views. In place of 
highlighting, one  could use opacity or binned colour scales to encode 
distances or ranks instead of the neighbouring points. We have begun 
implementing this brush in **liminal**,  using the  **FNN** or **RcppAnnoy**
packages for fast neighbourhood estimation on the server side, however there
are still technicalaties that need be resolved [@fnn-pkg; @rcpp-annoy-pkg]. 
Brush composition, such as 'and', 'or', or 'not' brushes, could be used to 
further investigate mismatches between the $k$-nn graphs estimated from both the embedding and tour views.

>>>>

We have shown in the case studies, that one to one linked brushing can
be used to identify distortions in the embedding, however we would
like extend this to one to many linked brushing, which would allow us to directly interrogate neighbourhood preservation. This form of brushing acts directly on a $k$-nearest neighbours ($k$-nn) graph computed from a
reference dataset: when a user brushes over a region in the embedding,
all the points that match the graphs edges are selected on the 
corresponding tour view. The reference data set for computing
nearest neighbours (for example a distance matrix, or the complete data matrix) 
can be independent of the tour or embedding views. In place of 
highlighting, one  could use opacity or binned colour scales to encode 
distances or ranks instead of the neighbouring points. We have begun 
implementing this brush in **liminal**,  using the  **FNN** or **RcppAnnoy**
packages for fast neighbourhood estimation on the server side, however there
are still technicalities that need be resolved [@fnn-pkg; @rcpp-annoy-pkg]. 
Brush composition, such as 'and', 'or', or 'not' brushes, could be used to 
further investigate mismatches between the $k$-nn graphs estimated from both the embedding and tour views.
